Measuring causal influence with\\ back-to-back regression: the linear case - supplementary material

### 2019-05-24 paper/supplementary.tex
Some typos

@@

@@

\clearpage
 \newpage
 \begin{theorem}[B2B consistency - general case]

     Consider the B2B model from Equation $Y = (XE + N)F$, $N$ centred and full rank noise.
     %
     If $F$ and $X$ are full-rank on $Img(E)$, then, the solution of B2B, $\hat H$, will minimise
     %
     $\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$and satisfy $E\hat H = \hat H$
\end{theorem}
\begin{proof}

 Let $\hat G$ and $\hat H$ be the solutions of the first and second regressions of B2B, we have
 \begin{align*}
    \hat G = \arg \min_G \mathbb{E}[\left \| YG - X \right \|^2] &=   \arg \min_G \mathbb{E}[\left \| X - (XE + N)FG \right\|^2]\\
                                                        &{}= \arg \min_G \mathbb{E}[\left \| X - XEFG\right\| ^2]  + \mathbb{E}[\left \| NFG\right \| ^2]\\
                                                        &{}= \arg \min_G \left \| X - XEFG\right\| ^2  + \left \| NFG\right \| ^2
     \label{eq:doublenorm}
\end{align*}
\begin{align*}
    \hat H = \arg \min_H \mathbb{E}[\| XH - Y \hat{G} \|^2] &=\arg  \min_H \mathbb{E}[\| XH - (XE + N)F \hat G \|^2] \\
    &=\arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2] + \mathbb{E}[\| NF\hat G \| ^2]\\
    &= \arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2]\\
    &= EF \hat G
\end{align*}

Let us prove that $EF\hat G = F\hat G$, that is, that the $d_x$ bottom rows of $F\hat G$ are zero.

 Let $Z=F^\dagger EF\hat G$. We have $FZ = FF^\dagger EF  \hat G= EF\hat G$ ($FF^\dagger E =E$ since $F$ is full rank on $Img(E)$). Since E is a contraction, we have $ \| NEF\hat G\|^2 \leq \| NF\hat G \|^2$. Therefore, 
 $$\left \| X - XEFZ\right \| ^2  + \left \| NFZ\right \| ^2 = \| X - XEF\hat G \| ^2  + \| NEF\hat G \| ^2 \leq \| X - XEF\hat G \| ^2  + \| NF\hat G \| ^2$$

@@

$Diag_k (\hat H) = Diag((\Sigma_{X1}+\Sigma_{N1})^{-1}\Sigma_{X1}) = Diag((I+\Sigma_{X1}^{-1}\Sigma_{N1})^{-1})$

In the absence of noise, we have $Diag(\hat H) = Diag(E)$. Else the k first diagonal elements of $\hat H$ are all positive, and bounded by $\frac{\sigma_{X_k}}{\sigma_{X_k} +\sigma_{N_1}}$ and $\frac{\sigma_{X_1}}{\sigma_{X_1} +\sigma_{N_k}}$, where  $\sigma_{X_1}$, $\sigma_{X_k}$, $\sigma_{N_1}$ and $\sigma_{N_k}$ denote the largest and smallest eigenvalues of $\Sigma_{X1}$ and $\Sigma_{N1}$.

The average value of the coefficients of $Diag(\hat H)$ is the trace of $\hat H$ divided by k. On average, since $X$ and $N$ are not correlated, $\Sigma_{X1}^{-1}\Sigma_{N1}$ should have a low condition numbers, and $\frac{Var(X)}{Var(X)+Var(N)}$ provide a good estimate of $\frac{1}{k}Tr(\hat H)$.

>>>>

\appendix

@@

\section{Additional Figures}

@@

\clearpage
\newpage

\section{Additional Theorems}

\setcounter{theorem}{1}
 \begin{theorem}[B2B consistency, general case]

     Consider the B2B model from Equation $Y = (XE + N)F$, with centered, full-rank noise $N$.
     %
     If $F$ and $X$ are full-rank on $\text{Img}(E)$, then, the solution of B2B, $\hat H$, will minimise
     %
     $\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$and satisfy $E\hat H = \hat H$
\end{theorem}
\begin{proof}

 Let $\hat G$ and $\hat H$ be the solutions of the first and second regressions of B2B, we have
 \begin{align*}
    \hat G = \arg \min_G \mathbb{E}[\left \| YG - X \right \|^2] &=   \arg \min_G \mathbb{E}[\left \| X - (XE + N)FG \right\|^2]\\
                                                        &{}= \arg \min_G \mathbb{E}[\left \| X - XEFG\right\| ^2]  + \mathbb{E}[\left \| NFG\right \| ^2]\\
                                                        &{}= \arg \min_G \left \| X - XEFG\right\| ^2  + \left \| NFG\right \| ^2,
     \label{eq:doublenorm}
\end{align*}
\begin{align*}
    \hat H = \arg \min_H \mathbb{E}[\| XH - Y \hat{G} \|^2] &=\arg  \min_H \mathbb{E}[\| XH - (XE + N)F \hat G \|^2] \\
    &=\arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2] + \mathbb{E}[\| NF\hat G \| ^2]\\
    &= \arg \min_H \mathbb{E}[\| X(H - EF \hat G) \| ^2]\\
    &= EF \hat G.
\end{align*}
%
First, we show that $EF\hat G = F\hat G$.
%
Let $Z=F^\dagger EF\hat G$. We have $FZ = FF^\dagger EF  \hat G= EF\hat G$.
Then, $FF^\dagger E =E$ since $F$ is full rank on $\text{Img}(E)$.
%
Since $E$ is a contraction, we have $ \| NEF\hat G\|^2 \leq \| NF\hat G \|^2$. Therefore, 
 $$\left \| X - XEFZ\right \| ^2  + \left \| NFZ\right \| ^2 = \| X - XEF\hat G \| ^2  + \| NEF\hat G \| ^2 \leq \| X - XEF\hat G \| ^2  + \| NF\hat G \| ^2.$$

@@

    $\text{Diag}_k (\hat H) = \text{Diag}((\Sigma_{X1}+\Sigma_{N1})^{-1}\Sigma_{X1}) = \text{Diag}((I+\Sigma_{X1}^{-1}\Sigma_{N1})^{-1})$

In the absence of noise, we have $\text{Diag}(\hat H) = \text{Diag}(E)$. Else the k first diagonal elements of $\hat H$ are all positive, and bounded by $\frac{\sigma_{X_k}}{\sigma_{X_k} +\sigma_{N_1}}$ and $\frac{\sigma_{X_1}}{\sigma_{X_1} +\sigma_{N_k}}$, where  $\sigma_{X_1}$, $\sigma_{X_k}$, $\sigma_{N_1}$ and $\sigma_{N_k}$ denote the largest and smallest eigenvalues of $\Sigma_{X1}$ and $\Sigma_{N1}$.

The average value of the coefficients of $\text{Diag}(\hat H)$ is the trace of $\hat H$ divided by k. On average, since $X$ and $N$ are not correlated, $\Sigma_{X1}^{-1}\Sigma_{N1}$ should have a low condition numbers, and $\frac{Var(X)}{Var(X)+Var(N)}$ provide a good estimate of $\frac{1}{k}Tr(\hat H)$.
